Dynamical modeling of tidal streams

### 2014-01-02 tex/ms.tex
streams: first draft pre-reading it over

%BOVY:NOTES:
% - We/I distinction
\documentclass[12pt,preprint]{aastex}
%\documentclass{emulateapj}
\usepackage{amssymb,amsmath}
%\usepackage{caption}
%\DeclareCaptionLabelSeparator{dotemdash}{.--- }
%\captionsetup[figure]{labelformat=simple, labelsep=dotemdash}
%\usepackage{color,hyperref}
% hypertex insanity
%\definecolor{linkcolor}{rgb}{0,0,0.25}
%\hypersetup{
%  colorlinks=true,        % false: boxed links; true: colored links
%  linkcolor=linkcolor,    % color of internal links
%  citecolor=linkcolor,    % color of links to bibliography
%  filecolor=linkcolor,    % color of file links
%  urlcolor=linkcolor      % color of external links
%}
\newcounter{address}
\setcounter{address}{1}
%\usepackage{sidecap}
\setlength{\emergencystretch}{2em}%No overflowing references
\newcommand{\ie}{i.e.}
\newcommand{\etal}{et al.}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\eg}{e.g.}
\newcommand{\eqnname}{equation}
\newcommand{\Eqnname}{Equation}
\newcommand{\equationname}{\eqnname}
\renewcommand{\tablename}{Table}
\renewcommand{\figurename}{Figure}
\newcommand{\figurenames}{\figurename~s}
\newcommand{\sectionname}{$\mathsection$}
\newcommand{\normal}{\ensuremath{\mathcal{N}}}
\newcommand{\flag}[1]{\texttt{\lowercase{#1}}}
\newcommand{\feh}{\ensuremath{[\mathrm{Fe/H}]}}
\newcommand{\afe}{\ensuremath{[\alpha\mathrm{/Fe}]}}
\newcommand{\logg}{log g}
\newcommand{\Ro}{\ensuremath{R_0}}

@@

This paper is outlined as follows. In \sectionname~\ref{sec:dynamics}
I briefly summarize the dynamics of tidal streams in terms of
action--angle coordinates. A generative model of a tidal stream in
frequency--angle coordinates is given in
\sectionname~\ref{sec:generative}. \sectionname~\ref{sec:track}
describes how to compute the stream's properties as a function of
angle along the stream, both in action--angle coordinates and in
poition--velocity coordinates. \sectionname~\ref{sec:mock} discusses
how to generate mock stream data using the generative model of a tidal
stream and in \sectionname~\ref{sec:pdf} I show how to calculate the
stream PDF for individual stream members as well as its
marginalization and convolution over missing and noisy directions. A
discussion and outlook is presented in
\sectionname~\ref{sec:discussion} and I conclude in
\sectionname~\ref{sec:conclusion}. In \appendixname~\ref{sec:aa} a
novel method for calculating actions, frequencies, and angles for any
orbit in any static gravitational potential is presented.

@@

\begin{figure}[t!]
  \includegraphics[width=0.5\textwidth,clip=]{gd1_evol_xz.ps}
  \caption{Mock tidal stream simulated as described in
    \sectionname~\ref{sec:sim} after $5.011\Gyr$ of evolution in
    Galactocentric $X$ and $Z$ coordinates. The grey dot shows the
    current position of the progenitor cluster and the dashed line
    shows its orbit. The solid line gives the stream track, calculated
    using the method described in \sectionname~\ref{sec:track}. The
    inset shows a zoomed-in version of part of the leading stream that
    more clearly shows the offset between the stream track and the
    progenitor orbit.}\label{fig:gd1_xz}
%python plot_stream.py ../tex/gd1_evol_xz.ps
\end{figure}

@@

The action and frequency distributions shown in
\figurename~\ref{fig:gd1_jt} combined with the knowledge that for
\emph{any} tidal stream the Hessian $(\partial^2 H / \partial \vecj
\partial \vecj)\large|_{\vecj^p}$ must be strongly dominated by a
single large eigenvalue leads me to propose that frequency--angle
space is a better coordinate system than action--angle space for
modeling tidal streams. That is, the distribution of actions in a
tidal stream has a complicated structure that is hard to capture using
a simple, analytic form (see also the discussion in
\citealt{Eyre11a}). Especially the ``bowtie'' structure in $(J_R,L_Z)$
and the overlapping leading- and trailing-arm distributions in
($J_R,J_Z$) are difficult to model with a simple distribution
function. The distribution of frequencies, however, is close-to one
dimensional, its direction of largest variance can be well-modeled as
a Gaussian (although below we will model it slightly differently), and
its parameters can all be easily estimated from the velocity
dispersion of the progenitor, the orbit of the progenitor, and the
gravitational potential, as discussed below. Therefore, we will model
tidal streams in frequency--angle space. Because the leading and
trailing arm are well-separated in frequency space, we model each arm
individually.

@@

The middle panel of \figurename~\ref{fig:gd1_times} shows the
frequency offsets between stream particles and the progenitor as a
function of strippng time, corresponding to ingredient (b) above. The
same burstiness that is apparent in the distribution of stripping
times shows up here. Particles stripped at pericenter typically have
larger frequency offsets than those lost at larger radii. This can be
described as saying that particles are \emph{removed} at pericenter,
but only \emph{peeled off} between pericenter passages. Again,
realistic models of tidal streams need to take this detailed structure
into account.

@@

Because we expect the typical $\Delta\veco$ to scale with the velocity
dispersion of the progenitor, I conjecture that a constant
$\mu_\Omega$ can be used for modeling cold tidal streams of any
(small-ish) velocity dispersion. Further modeling of simulated
streams, however, is necessary to check this and to find the optimizal
value of $\mu_\Omega$. Fixing $\mu_\Omega$ leaves a single free
parameter \sigv\ for modeling the frequency offset distribution of a
tidal stream. In modeling to observed data, this parameter needs to be
fit and we expect it to be proportional to the velocity dispersion of
the progenitor, although this proportionality needs to be checked more
carefully (but see Figure 3 in \citealt{Sanders13a}, which shows that
the size of the frequency distribution scales as mass$^{1/3}$, or
approximately as the velocity dispersion). Similarly, the initial
angle distribution, modeled here as an isotropic Gaussian, has a
charactaristic spread $\sigma_{\theta}$ which I set here to $0.003$
($=\sigv/[122\kms]$), and we likewise expect this spread to scale with
the velocity dispersion of the progenitor.

@@

\figurename~\ref{fig:track} illustrates how I linearize the
$(\vecx,\vecv) \rightarrow (\veco,\veca)$ transformation along the
progenitor orbit and use it to propagate the track of the stream to
position--velocity coordinates and to observable quantities
$(l,b,D,\vlos,\pmll,\pmbb)$. The top panel shows the orbit of the
progenitor and the mean stream track in frequency--angle
coordinates. Only the parallel direction is shown here; the mean
stream track coincides with the progenitor orbit in the perpendicular
direction. At a small number of points along the progenitor orbit,
chosen here to span $1.5$ radians, we calculate the Jacobian $\partial
(\veco,\veca) / \partial (\vecx,\vecv)$. In this Figure this is done
at four points, but for all other figures and calculations in this
paper eleven points are used. Each Jacobian calculation requires seven
frequency--angle calcuations, such that the total number of such
computations is less than $100$ to adequately model the stream (each
computation involves a single orbit integration, see
\appendixname~\ref{sec:aa}).

@@

For the fiducial stream model of \sectionname~\ref{sec:fidmodel} the
procedure for producing mock data is simplified to sampling a small
number of easy-to-sample distributions. First, $t_s$ is drawn
uniformly between $0$ and $t_d$. Second, $\Delta \operp$ is drawn from
the Gaussian distribution of $\Delta \operp$. The distribution of
$\Delta \opar$ is not Gaussian, as it is a Gaussian in $|\Delta
\opar|$ multiplied with $|\Delta \opar|$ over the range $\Delta \opar
> 0$. This distribution is log-concave, that is, the derivative of its
logarithm is negative everywhere, and therefore it can be efficienly
sampled using adaptive-rejection sampling \citep{Gilks92a}. The sign
of $\Delta \opar$ is determined by whether we are generating a leading
or trailing stream. Third, initial angle offsets are drawn from the
Gaussian distribution of such offsets. The fourth step is as above.

@@

\figurename~\ref{fig:gd1_mock} shows mock data generated for the model
of the simulated stream described in
\sectionname~\ref{sec:fidmodel}. The mock data in this Figure has been
generated $1.96\Gyr$ before the final snapshot of the simulation (that
is, $1.96\Gyr$ before \figurename~\ref{fig:gd1_xz}). The model
from \sectionname~\ref{sec:fidmodel}, which was for the final
snapshot, was adjusted to this earlier time by backward orbit
integration of the model progenitor and by revising the disruption
time downward by $1.96\Gyr$. No other changes were made to the
model. The right panel of \figurename~\ref{fig:gd1_mock} shows the
mock data generated from the model in $(X,Z)$, while the left panel
shows the simulated stream data at this time. Overall, the
distribution of the mock and simulated data are similar.

@@

The dynamical framework proposed here can be used in situations where
fitting a stream track is insufficient, for exampe, when high-quality
data in some dimensions on individual stars are available. Many of the
approaches in the literature do not require a model progenitor
position. This includes methods that attempt to minimize the spread of
orbital energies or actions in the stream (\eg,
\citealt{Binney08a,Penarrubia12a}, Sanderson \etal, 2014, in
preparation) as well as the method of \citet{Sanders13b}, which
constrains the potential by requiring angle differences in the stream
to lie along the same direction as frequency differences (see
\equationname~\ref{eq:pdf} and subsequent discussion). These methods
require good measurements of the phase--space coordinates of stream
members, or at the very least, \emph{some} measurement of all of the
six phase--space coordinates (although the latter can be relaxed when
excellent measurements of the line-of-sight velocity along the stream
are available; \citealt{Binney08a}). If such high-quality data are
present, these methods can provide good initial guesses for the PDF of
potential parameters that fit a stream, which can subsequently be used
in a more thorough exploration of the PDF and structure of the stream
using the framework of \sectionname~\ref{sec:pdf}. That these methods
work also goes to show that the progenitor parameters---phase--space
position, $\sigv$ and $t_d$---are not that important for the
constraints on the gravitational potential.

@@

In our stream model the distribution of $\Delta \veco$ is constant
with time, such that stream members on average are removed to the same
orbit. However, the finite disruption time means that for stream
members to have reached the largest angle offsets from the progenitor,
they have to have been removed onto an orbit that is separated from
the progenitor's by a larger $\Delta \veco$ than average. In our model
this is explicitly expressed by the distribution $p(\Delta
\opar|\Delta \apar)$, calculated in \equationname~(\ref{eq:poparapar})
and shown in the top left panel of
\figurename~\ref{fig:gd1_apar}. Thus, the average orbit changes as a
function of positon along the stream, whether expressed as $\Delta
\apar$, $X$, or $l$, and a single-orbit approximation is necessarily
wrong over the full length of the stream.

@@

The framework allows for the proper analysis of data on tidal streams
by taking into account all the relevant dynamical effects, most
notably the stream--single-orbit offset. As such, it should be
immediately useful for the proper analysis of data on the GD-1
\citep{Koposov10a}, Orphan \citep[\eg,][]{Sesar13a}, Pal 5 and othter
streams. The ability to quickly marginalize the model over unobserved
dimensions will also prove highly useful when Gaia data for these
streams becomes available in the near future. The novel framework also
provides a straightforward way to generate mock stream data that is
useful for detecting anomalies in the structure of tidal streams, such
as those that can be caused by fly-bys of dark--mater subhalos.

@@

The Fox \& Binney method consists of averaging \eqnname~(\ref{eq:jjt})
over the auxiliary angle along the path of the orbit, such that the
mapping defined by the $S_{\vecn}$ is onto orbital tori of the target
Hamiltonian, the oscillating cosine behavior cancels, and the target
action is obtained as the auxiliary-angle-averaged auxilary
action. That is,
\begin{equation}\label{eq:jjtint}
  \int \dd \theta_i^A J_i^A = \int \dd \theta_i^A J_i + 2 \sum_{\vecn > 0} \vecn\,S_{\vecn}(J_i)\int \dd \theta_i^A\cos(\vecn\cdot\veca^A)\,,
\end{equation}
where the integral is over auxiliary angles calculated on the path of
the orbit in the target potential. The factor $\vecn\,S_{\vecn}(J_i)$
comes out of the integral as the target actions are conserved along
the orbit. In this Equation, we integrate over the angle coordinate
$\theta_i^A$ associated with the action $J_i$ that we want to compute
(that is, $\theta_R$ for $J_R$ and so on). This is because of the
factor $\vecn$: the condition $\vecn > 0$ means that when calculating
$J_R$ there are terms with $\vecn = (n_R,0,0)$, which would not cancel
according to the argument below when integrating over $\theta_Z$. For
non-resonant orbits and after a long-enough orbit integration, the
angles can be considered to be independent (that is, the orbit ``fills
its torus''), such that we can write (for example, for $\theta_R$)
\begin{equation}
\begin{split}
  \int \dd \theta_R^A\cos(\vecn\cdot\veca^A) & = \mathrm{Re}\left[\int \dd \theta_R^A\,\exp(-i\vecn\cdot\veca^A)\right]\,, \\
  & \approx \mathrm{Re}\left[\exp(-i\,n_\phi\,\theta^A_\phi-i\,n_Z\,\theta^A_Z)\int \dd \theta_R^A\,\exp(-i\,n_R\,\theta^A_R)\right]\,.
\end{split}
\end{equation}
As long as the auxiliary angle $\theta_R$ goes through the full range
$0$ to $2\pi$ along the orbit, the oscillatory behavior integrates to
zero and \eqnname~(\ref{eq:jjtint}) simplifies to
\begin{equation}\label{eq:aAIJ}
  J_i = \frac{\int \dd \theta_i^A J_i^A}{\int \dd \theta_i^A}\,,
\end{equation}
that is, the target action is obtained by averaging the auxiliary
action over the auxiliary angle.

>>>>

\documentclass[12pt,preprint]{aastex}
%\documentclass{emulateapj}
\usepackage{amssymb,amsmath}
%\usepackage{caption}
%\DeclareCaptionLabelSeparator{dotemdash}{.--- }
%\captionsetup[figure]{labelformat=simple, labelsep=dotemdash}
%\usepackage{color,hyperref}
% hypertex insanity
%\definecolor{linkcolor}{rgb}{0,0,0.25}
%\hypersetup{
%  colorlinks=true,        % false: boxed links; true: colored links
%  linkcolor=linkcolor,    % color of internal links
%  citecolor=linkcolor,    % color of links to bibliography
%  filecolor=linkcolor,    % color of file links
%  urlcolor=linkcolor      % color of external links
%}
\newcounter{address}
\setcounter{address}{1}
%\usepackage{sidecap}
\setlength{\emergencystretch}{2em}%No overflowing references
\newcommand{\ie}{i.e.}
\newcommand{\etal}{et al.}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\eg}{e.g.}
\newcommand{\eqnname}{equation}
\newcommand{\Eqnname}{Equation}
\newcommand{\equationname}{\eqnname}
\renewcommand{\tablename}{Table}
\renewcommand{\figurename}{Figure}
\newcommand{\figurenames}{\figurename~s}
\newcommand{\sectionname}{$\mathsection$}
\newcommand{\normal}{\ensuremath{\mathcal{N}}}
\newcommand{\flag}[1]{\texttt{\lowercase{#1}}}
\newcommand{\feh}{\ensuremath{[\mathrm{Fe/H}]}}
\newcommand{\afe}{\ensuremath{[\alpha\mathrm{/Fe}]}}
\newcommand{\logg}{log g}
\newcommand{\Ro}{\ensuremath{R_0}}

@@

This paper is outlined as follows. In \sectionname~\ref{sec:dynamics}
I briefly summarize the dynamics of tidal streams in terms of
action--angle coordinates. A generative model of a tidal stream in
frequency--angle coordinates is given in
\sectionname~\ref{sec:generative}. \sectionname~\ref{sec:track}
describes how to compute the stream's properties as a function of
angle along the stream, both in action--angle coordinates and in
position--velocity coordinates. \sectionname~\ref{sec:mock} discusses
how to generate mock stream data using the generative model of a tidal
stream and in \sectionname~\ref{sec:pdf} I show how to calculate the
stream PDF for individual stream members as well as its
marginalization and convolution over missing and noisy directions. A
discussion and outlook is presented in
\sectionname~\ref{sec:discussion} and I conclude in
\sectionname~\ref{sec:conclusion}. In \appendixname~\ref{sec:aa} a
novel method for calculating actions, frequencies, and angles for any
orbit in any static gravitational potential is presented.

@@

\begin{figure}[t!]
  \includegraphics[width=0.5\textwidth,clip=]{gd1_evol_xz.ps}
  \caption{Mock tidal stream simulated as described in
    \sectionname~\ref{sec:sim} after $5.011\Gyr$ of evolution in
    Galactocentric $X$ and $Z$ coordinates. The gray dot shows the
    current position of the progenitor cluster and the dashed line
    shows its orbit. The solid line gives the stream track, calculated
    using the method described in \sectionname~\ref{sec:track}. The
    inset shows a zoomed-in version of part of the leading stream that
    more clearly shows the offset between the stream track and the
    progenitor orbit.}\label{fig:gd1_xz}
%python plot_stream.py ../tex/gd1_evol_xz.ps
\end{figure}

@@

The action and frequency distributions shown in
\figurename~\ref{fig:gd1_jt} combined with the knowledge that for
\emph{any} tidal stream the Hessian $(\partial^2 H / \partial \vecj
\partial \vecj)\large|_{\vecj^p}$ must be strongly dominated by a
single large eigenvalue leads me to propose that frequency--angle
space is a better coordinate system than action--angle space for
modeling tidal streams. That is, the distribution of actions in a
tidal stream has a complicated structure that is hard to capture using
a simple, analytic form (see also the discussion in
\citealt{Eyre11a}). Especially the ``bow tie'' structure in $(J_R,L_Z)$
and the overlapping leading- and trailing-arm distributions in
($J_R,J_Z$) are difficult to model with a simple distribution
function. The distribution of frequencies, however, is close-to one
dimensional, its direction of largest variance can be well-modeled as
a Gaussian (although below we will model it slightly differently), and
its parameters can all be easily estimated from the velocity
dispersion of the progenitor, the orbit of the progenitor, and the
gravitational potential, as discussed below. Therefore, we will model
tidal streams in frequency--angle space. Because the leading and
trailing arm are well-separated in frequency space, we model each arm
individually.

@@

The middle panel of \figurename~\ref{fig:gd1_times} shows the
frequency offsets between stream particles and the progenitor as a
function of stripping time, corresponding to ingredient (b) above. The
same burstiness that is apparent in the distribution of stripping
times shows up here. Particles stripped at pericenter typically have
larger frequency offsets than those lost at larger radii. This can be
described as saying that particles are \emph{removed} at pericenter,
but only \emph{peeled off} between pericenter passages. Again,
realistic models of tidal streams need to take this detailed structure
into account.

@@

Because we expect the typical $\Delta\veco$ to scale with the velocity
dispersion of the progenitor, I conjecture that a constant
$\mu_\Omega$ can be used for modeling cold tidal streams of any
(small-ish) velocity dispersion. Further modeling of simulated
streams, however, is necessary to check this and to find the optimal
value of $\mu_\Omega$. Fixing $\mu_\Omega$ leaves a single free
parameter \sigv\ for modeling the frequency offset distribution of a
tidal stream. In modeling to observed data, this parameter needs to be
fit and we expect it to be proportional to the velocity dispersion of
the progenitor, although this proportionality needs to be checked more
carefully (but see Figure 3 in \citealt{Sanders13a}, which shows that
the size of the frequency distribution scales as mass$^{1/3}$, or
approximately as the velocity dispersion). Similarly, the initial
angle distribution, modeled here as an isotropic Gaussian, has a
characteristic spread $\sigma_{\theta}$ which I set here to $0.003$
($=\sigv/[122\kms]$), and we likewise expect this spread to scale with
the velocity dispersion of the progenitor.

@@

\figurename~\ref{fig:track} illustrates how I linearize the
$(\vecx,\vecv) \rightarrow (\veco,\veca)$ transformation along the
progenitor orbit and use it to propagate the track of the stream to
position--velocity coordinates and to observable quantities
$(l,b,D,\vlos,\pmll,\pmbb)$. The top panel shows the orbit of the
progenitor and the mean stream track in frequency--angle
coordinates. Only the parallel direction is shown here; the mean
stream track coincides with the progenitor orbit in the perpendicular
direction. At a small number of points along the progenitor orbit,
chosen here to span $1.5$ radians, we calculate the Jacobian $\partial
(\veco,\veca) / \partial (\vecx,\vecv)$. In this Figure this is done
at four points, but for all other figures and calculations in this
paper eleven points are used. Each Jacobian calculation requires seven
frequency--angle calculations, such that the total number of such
computations is less than $100$ to adequately model the stream (each
computation involves a single orbit integration, see
\appendixname~\ref{sec:aa}).

@@

For the fiducial stream model of \sectionname~\ref{sec:fidmodel} the
procedure for producing mock data is simplified to sampling a small
number of easy-to-sample distributions. First, $t_s$ is drawn
uniformly between $0$ and $t_d$. Second, $\Delta \operp$ is drawn from
the Gaussian distribution of $\Delta \operp$. The distribution of
$\Delta \opar$ is not Gaussian, as it is a Gaussian in $|\Delta
\opar|$ multiplied with $|\Delta \opar|$ over the range $\Delta \opar
> 0$. This distribution is log-concave, that is, the derivative of its
logarithm is negative everywhere, and therefore it can be efficiently
sampled using adaptive-rejection sampling \citep{Gilks92a}. The sign
of $\Delta \opar$ is determined by whether we are generating a leading
or trailing stream. Third, initial angle offsets are drawn from the
Gaussian distribution of such offsets. The fourth step is as above.

@@

\figurename~\ref{fig:gd1_mock} shows mock data generated for the model
of the simulated stream described in
\sectionname~\ref{sec:fidmodel}. The mock data in this Figure have
been generated $1.96\Gyr$ before the final snapshot of the simulation
(that is, $1.96\Gyr$ before \figurename~\ref{fig:gd1_xz}). The model
from \sectionname~\ref{sec:fidmodel}, which was for the final
snapshot, was adjusted to this earlier time by backward orbit
integration of the model progenitor and by revising the disruption
time downward by $1.96\Gyr$. No other changes were made to the
model. The right panel of \figurename~\ref{fig:gd1_mock} shows the
mock data generated from the model in $(X,Z)$, while the left panel
shows the simulated stream data at this time. Overall, the
distribution of the mock and simulated data are similar.

@@

The dynamical framework proposed here can be used in situations where
fitting a stream track is insufficient, for example, when high-quality
data in some dimensions on individual stars are available. Many of the
approaches in the literature do not require a model progenitor
position. This includes methods that attempt to minimize the spread of
orbital energies or actions in the stream (\eg,
\citealt{Binney08a,Penarrubia12a}, Sanderson \etal, 2014, in
preparation) as well as the method of \citet{Sanders13b}, which
constrains the potential by requiring angle differences in the stream
to lie along the same direction as frequency differences (see
\equationname~\ref{eq:pdf} and subsequent discussion). These methods
require good measurements of the phase--space coordinates of stream
members, or at the very least, \emph{some} measurement of all of the
six phase--space coordinates (although the latter can be relaxed when
excellent measurements of the line-of-sight velocity along the stream
are available; \citealt{Binney08a}). If such high-quality data are
present, these methods can provide good initial guesses for the PDF of
potential parameters that fit a stream, which can subsequently be used
in a more thorough exploration of the PDF and structure of the stream
using the framework of \sectionname~\ref{sec:pdf}. That these methods
work also goes to show that the progenitor parameters---phase--space
position, $\sigv$ and $t_d$---are not that important for the
constraints on the gravitational potential.

@@

In our stream model the distribution of $\Delta \veco$ is constant
with time, such that stream members on average are removed to the same
orbit. However, the finite disruption time means that for stream
members to have reached the largest angle offsets from the progenitor,
they have to have been removed onto an orbit that is separated from
the progenitor's by a larger $\Delta \veco$ than average. In our model
this is explicitly expressed by the distribution $p(\Delta
\opar|\Delta \apar)$, calculated in \equationname~(\ref{eq:poparapar})
and shown in the top left panel of
\figurename~\ref{fig:gd1_apar}. Thus, the average orbit changes as a
function of position along the stream, whether expressed as $\Delta
\apar$, $X$, or $l$, and a single-orbit approximation is necessarily
wrong over the full length of the stream.

@@

The framework allows for the proper analysis of data on tidal streams
by taking into account all the relevant dynamical effects, most
notably the stream--single-orbit offset. As such, it should be
immediately useful for the proper analysis of data on the GD-1
\citep{Koposov10a}, Orphan \citep[\eg,][]{Sesar13a}, Pal 5 and other
streams. The ability to quickly marginalize the model over unobserved
dimensions will also prove highly useful when Gaia data for these
streams become available in the near future. The novel framework also
provides a straightforward way to generate mock stream data that is
useful for detecting anomalies in the structure of tidal streams, such
as those that can be caused by fly-bys of dark--mater subhalos.

@@

The Fox \& Binney method consists of averaging \eqnname~(\ref{eq:jjt})
over the auxiliary angle along the path of the orbit, such that the
mapping defined by the $S_{\vecn}$ is onto orbital tori of the target
Hamiltonian, the oscillating cosine behavior cancels, and the target
action is obtained as the auxiliary-angle-averaged auxiliary
action. That is,
\begin{equation}\label{eq:jjtint}
  \int \dd \theta_i^A J_i^A = \int \dd \theta_i^A J_i + 2 \sum_{\vecn > 0} \vecn\,S_{\vecn}(J_i)\int \dd \theta_i^A\cos(\vecn\cdot\veca^A)\,,
\end{equation}
where the integral is over auxiliary angles calculated on the path of
the orbit in the target potential. The factor $\vecn\,S_{\vecn}(J_i)$
comes out of the integral as the target actions are conserved along
the orbit. In this Equation, we integrate over the angle coordinate
$\theta_i^A$ associated with the action $J_i$ that we want to compute
(that is, $\theta_R$ for $J_R$ and so on). This is because of the
factor $\vecn$: the condition $\vecn > 0$ means that when calculating
$J_R$ there are terms with $\vecn = (n_R,0,0)$, which would not cancel
according to the argument below when integrating over $\theta_Z$. For
non-resonant orbits and after a long-enough orbit integration, the
angles can be considered to be independent (that is, the orbit ``fills
its torus''), such that we can write (for example, for $\theta_R$)
\begin{equation}
\begin{split}
  \int \dd \theta_R^A\cos(\vecn\cdot\veca^A) & = \mathrm{Re}\left[\int \dd \theta_R^A\,\exp(-i\vecn\cdot\veca^A)\right]\,, \\
  & \approx \mathrm{Re}\left[\exp(-i\,n_\phi\,\theta^A_\phi-i\,n_Z\,\theta^A_Z)\int \dd \theta_R^A\,\exp(-i\,n_R\,\theta^A_R)\right]\,.
\end{split}
\end{equation}
As long as the auxiliary angle $\theta_R$ goes through the full range
$0$ to $2\pi$ along the orbit, the oscillatory behavior integrates to
zero and \eqnname~(\ref{eq:jjtint}) simplifies to
\begin{equation}\label{eq:aAIJ}
  J_i = \frac{\int \dd \theta_i^A J_i^A}{\int \dd \theta_i^A}\,,
\end{equation}
that is, the target action is obtained by averaging the auxiliary
action over the auxiliary angle.
